---
date: "2023-11-08"
title: "Class G, Mocap data"
author: "Luke Ring / zeyus + <YOUR NAME HERE>"
output: html_document
---

# Class G, Mocap data analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# to make things easy, we will use pacman.
# Please install it if you don't have it already
# with install.packages("pacman")

# packages to load
pacman::p_load(
  "XML",
  "tidyverse",
  "fs",
  "assertthat",
  "stringi",
  "dtw",
  "RTransferEntropy",
  "signal",
  "conflicted",
  "Rcpp",
  "future"
)
```

Some setup...You will need to make sure the data_dir correctly points to where your .tsv files are.
e.g., if your files are on your Desktop (please don't) in the folder `tsvs`, you would use:

```{r, eval=FALSE}
data_dir <- path_home() %>%
  path(
    "Desktop",
    "tsvs"
  )

```

Now you can update the `data_dir` variable below.

**Note:** The `path_home()` function will just give you your home directory (e.g., `/Users/me` or `C:\Users\me`). The `path()` function will join together the different parts of the path.

```{r 'More setup', include=FALSE}

# make sure we are in the G_Class_6 directory
wd <- getwd()
if (basename(wd) != "G_Class_6") {
  setwd("./G_Class_6")
}

# set the directory for the data
# note, path_home() will return the path to your home directory
# your home directory is the one that contains your documents, downloads, etc.
data_dir <- path_home() %>%
  path(
    "Documents",
    "PercActClass",
    "labeled_data",
    "tsvs"
  )

# load the mocap_utils.R file with helper functions
source("mocap_utils.R")
```

## Load a single TSV file

Choose which file you want to use, e.g. `group0_LeadFollow.tsv`.

```{r 'Load data from a mocap session'}
# set file name
tsv_file_name <- "<FILE NAME HERE>"
# load the file
tsv_file_path <- path(data_dir, tsv_file_name)
mocap_data <- process_qtm_tsv(
  tsv_file_path,
  only_data = TRUE,
  long_format = TRUE
)

# take a look at the data
mocap_data
```

## Extract marker data for both subjects

Now that the data is ready to use, you need to pick a marker and axis for both subjects, for example,
you may want to look at `hand_right` on the `z` axis. You could also consider comparing one suject's
left hand with the other's right hand or anything else you want to try.

```{r 'Extract marker data'}

subj_a_data <- mocap_data %>%
  dplyr::filter(
    subject == "A",
    marker == "hand_right",
  )

subj_b_data <- mocap_data %>%
  dplyr::filter(
    subject == "B",
    marker == "hand_right",
  )

```

## Plot the time series

Take a look at the data, plot them, you might see some gaps, or you might want to start it a bit later or end it early.

```{r 'Plot the time series'}

# plot the data
subj_a_data %>%
  ggplot(aes(x = index, y = z)) +
  geom_line() +
  labs(
    title = "Subject A",
    x = "Index (sample)",
    y = "z Position (mm)"
  )

subj_b_data %>%
  ggplot(aes(x = index, y = z)) +
  geom_line() +
  labs(
    title = "Subject B",
    x = "Index (sample)",
    y = "z Position (mm)"
  )

```

## Remove tail and end

You probably have somethin at the start you don't want to include (t-pose), maybe at the end as well.
Try removing the start and end by the sample index and plot it again to see what it looks like after.

```{r 'Remove start and end'}

# the index that you want to start from
start_index <- 0 # put a number here based on your plot
# the index that you want to end at
end_index <- 1000 # put a number here based on your plot

# filter the data
subj_a_data <- subj_a_data %>%
  dplyr::filter(
    index >= start_index, # only keep rows where index is greater than or equal to start_index
    index <= end_index  # only keep rows where index is less than or equal to end_index
  )
subj_b_data <- subj_b_data %>%
  dplyr::filter(
    index >= start_index,
    index <= end_index
  )

# plot it again and see if it looks better
```

## Get just the time series for z-axis

We only want to work with one axis, so let's just get the time-series for both subjects.

```{r 'Get just the time series for z-axis'}
z_a <- subj_a_data$z
z_b <- subj_b_data$z
```

## Fill gaps

Most likely there are a few missing spots in the data, don't despair,
I've provided a helper function to do a linear gap fill. Let's do that now.

```{r 'Fill gaps'}

# fill the gaps
z_a_filled <- gap_fill_linear(z_a)
z_b_filled <- gap_fill_linear(z_b)

# plot it again and see if it looks better
plot(z_a_filled)
plot(z_b_filled)
```

## Run cross-correlation analysis

Let's get to it. This is a Cross correlation analysis. This works by correlating the time series of one marker
for subject A with the time series of the same marker for subject B. This can be useful for seeing if there is
any lag between the two subjects, or if there is any correlation between the two subjects.

What do you think lag means? What does correlated in this context mean?

Try and think about this in the context of movement, and then include the fact that the data is represented
as numbers along an axis, will this affect anything?

`ccf` is a function from the `stats` package, you can read more about it here:
https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/acf


```{r 'Run cross-correlation analysis'}

ccf_result <- ccf(
  z_a_filled,
  z_b_filled,
  lag.max = 1000,
  plot = FALSE,
  na.action = na.omit
)

# plot the result
plot(ccf_result)
```

## Run Dynamic Time warping

This is a way of comparing two time series that is used in speech recognition,
music information retrieval, neuroscience, and many other fields.
The main function that runs the dynamic time warping is `dtw` from the `dtw` package.
You can find more information here https://dtw.r-forge.r-project.org/

Try changing the windowing method, etc to see how they compare. Try the different plot types as well.

*Warning: This will take some time, DTW is very CPU intensive.*

You may want to consider downsampling the data.

```{r 'Run Dynamic Time warping'}

# run the dtw
dtw_result <- dtw(
  z_a_filled,
  z_b_filled,
  window.type = "none",
  keep.internals = TRUE
)

# plot the result
plot(
  dtw_result,
  type = "alignment"
)
plot(
  dtw_result,
  type = "density"
)
```

The main thing of interest that we get out of DTW is the distance between the two time series.
This is either the total distance between the two, or a normalized distance (this is the one you would use
if you were to compare multiple pairs of time series, e.g. different sujects, different conditions).

```{r 'Get the distance'}
# get the distance
dtw_result$distance
dtw_result$normalizedDistance
```

## Run Transfer Entropy

This is a way of measuring the transfer of information between two time series. In the context 
of the movement of two subjects, this can be useful for seeing if one subject is leading the other, for example.

Can you think of any other ways this might be useful?

There are various R packages for doing transfer entropy analysis, in this case we use the `RTransferEntropy` package.
https://cran.r-project.org/web/packages/RTransferEntropy/vignettes/transfer-entropy.html

```{r 'Run Transfer Entropy'}	
# run the transfer entropy
plan(multisession)
te_result <- transfer_entropy(
  z_a_filled,
  z_b_filled
)

te_result
```
